Classical Planning Algorithms ¡ on the Atari Video Games

### 2014-12-01 icaps15-paper/paper.tex
added ack
Fixed numbers

The number of nodes \emph{generated} by $\IW(1)$ is $n \times D \times b$ in the worst case,
where $n$ is the number of problem variables, $D$ is the size of their domains, and $b$ is the number of
actions per  state.  This same number in a  breadth-first search is not linear in $n$
but exponential. For the Atari games,  $n=128$, $D=256$, and $b=18$, so that the product is
equal to $589,824$, which is large but feasible. On the other hand, the number
of nodes generated by $\IW(2)$ in the worst case is  $(n \times D)^2 \times b$, which is
equal to $1,073,741,824$ which is too large, forcing us to consider only a tiny fraction of
such states. For classical planning problems, the growth in the number of nodes from
$\IW(1)$ to $\IW(2)$ is not that large, as the variables are boolean. Indeed, we could have
taken the state vector for the Atari games as a vector of 1024 boolean variables,
and apply these algorithms to that representation. The results would be different.
In such a case, $\IW(1)$ and $\IW(2)$  would generate $n' \times D' \times b$ and
$(n' \times D')^2 \times b$ states, which for $n'=1024$ and $D=2$
represent  $36,864$ and $7,5497,472$ states respectively.
These are feasible numbers, but we haven't used  this representation 
under the assumption that the correlations among the bits in each one of
the 128 words of the state vectors are meaningful.
In summary, from the basic \IW\ algorithm we are testing only the first
call $\IW(1)$.

@@

\input{games-table}
We tested \IW(1) and \BFS\ over \numgames\  of the $55$ games 
considered in \cite{bellemare:jair2013}, from now on  abbreviated as BNVB.\footnote{
We left out  \textsc{Skiing} as the reported figures apparently use a  different reward structure.}
%% check
% \footnote{
% Entries marked as  ``--'' in the table are   still running at the time of submission. 
% Each run can take up to $24$ hours of CPU time and there are $10$ runs per game and algorithm. 
% Eventually all of  the  $55$ games will  be included in the evaluation.}
Table~\ref{performance} shows the performance of $\IW(1)$ and \BFS\
in comparison  with breadth-first search (\BRFS) and \UCT.
Videos of selected games played by $\IW(1)$, \BFS, and \UCT\ can be 
seen in Youtube.\footnote{\url{http://www.youtube.com/playlist?list=PLXpQcXUQ_CwenUazUivhXyYvjuS6KQOI0}.}
The discount factor used by  all the algorithms is $\gamma=0.995$.
The scores reported for \BRFS\ and \UCT\ are taken from BNVB.  
Our experimental setup  follows theirs except that a maximum budget of  $150,000$ simulated frames 
is applied to $\IW(1)$, \BFS, and \UCT. \UCT\ uses this budget by running
 $500$   rollouts of depth $300$. % \footnote{In \UCT,  $500$ base policy rollouts
% of depth $300$ is equivalent to a budget of $150,000$ simulated frames.}.
The bound on the number of simulated frames is like a bound on
lookahead time, as most of the time in the lookahead  is spent
in calls to the emulator for computing the next RAM state. 
This is why the average time per action is similar to all the algorithms
except $\IW(1)$, that due to its aggressive pruning does not always use
the full budget and takes less time per action on average.

@@

\noindent \textbf{Acknowledgments.} We thank the people involved in the creation and maintenance of the 
Arcade Learning Environment  that have made this work possible.

>>>>

The number of nodes \emph{generated} by $\IW(1)$ is $n \times D \times b$ in the worst case,
where $n$ is the number of problem variables, $D$ is the size of their domains, and $b$ is the number of
actions per  state.  This same number in a  breadth-first search is not linear in $n$
but exponential. For the Atari games,  $n=128$, $D=256$, and $b=18$, so that the product is
equal to $589,824$, which is large but feasible. On the other hand, the number
of nodes generated by $\IW(2)$ in the worst case is  $(n \times D)^2 \times b$, which is
equal to $19,327,352,832$ which is too large, forcing us to consider only a tiny fraction of
such states. For classical planning problems, the growth in the number of nodes from
$\IW(1)$ to $\IW(2)$ is not that large, as the variables are boolean. Indeed, we could have
taken the state vector for the Atari games as a vector of 1024 boolean variables,
and apply these algorithms to that representation. The results would be different.
In such a case, $\IW(1)$ and $\IW(2)$  would generate $n' \times D' \times b$ and
$(n' \times D')^2 \times b$ states, which for $n'=1024$ and $D=2$
represent  $36,864$ and $75,497,472$ states respectively.
These are feasible numbers, but we haven't used  this representation 
under the assumption that the correlations among the bits in each one of
the 128 words of the state vectors are meaningful.
In summary, from the basic \IW\ algorithm we are testing only the first
call $\IW(1)$.

@@

\begin{table*}
\centering
% \addtolength{\tabcolsep}{-2pt}
% {\scriptsize
{\small
\begin{tabular}{@{}l|rr|rr|r|r@{}}
%% table header
% \hline\\[-3ex]
%\hline
\multicolumn{1}{c|}{}& \multicolumn{2}{c|}{$\IW(1)$}& \multicolumn{2}{c|}{\BFS} & \multicolumn{1}{c|}{\BRFS} &  \multicolumn{1}{c}{\UCT}
% \\[.2ex]
\\
% \hline\\\\[-3ex]
\hline
%% table body
\multicolumn{1}{c|}{Game}  &
% \multicolumn{1}{c|}{Avg. Score} &
% \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c|}{Avg. Time}  &
% \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c|}{Avg. Time}  &
% \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c|}{Avg. Time}  &
% \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c}{Avg. Time}\\[.2ex]
 \multicolumn{1}{c}{Score} & \multicolumn{1}{c|}{Time}  &
 \multicolumn{1}{c}{Score} & \multicolumn{1}{c|}{Time}  &
 \multicolumn{1}{c|}{Score} &
 \multicolumn{1}{c}{Score} 
% \\[.2ex]
\\[.2ex]
% \hline\\
\hline
\Alien & \mc{1}{\textbf{25634}} & 81 & \mc{1}{12252} & 81 & 784 & 7785 \\ \hline
\Amidar & \mc{1}{\textbf{1377}} & 28 &\mc{1}{ 1090} & 37 & 5 & 180 \\ \hline
\Assault & 953 & 18 & 827 & 25 & 414 & \textbf{1512} \\ \hline
\Asterix & 153400 & 24 & 77200 & 27 & 2136 & \textbf{290700} \\ \hline
\Asteroids & \mc{1}{\textbf{51338}} & 66 & \mc{1}{22168} & 65 & 3127 & 4661 \\ \hline
\Atlantis & 159420 & 13 & 154180 & 71 & 30460 & \textbf{193858} \\ \hline
\BankHeist & \mc{1}{\textbf{717}} & 39 & 362 & 64 & 22 & 498 \\ \hline
\BattleZone & 11600 & 86 &\mc{1}{ \textbf{330800}} & 87 & 6313 & 70333 \\ \hline
\BeamRider &\mc{1}{ 9108} & 23 & \mc{1}{\textbf{9298}} & 29 & 694 & 6625 \\ \hline
\Berzerk & \mc{1}{\textbf{2096}} & 58 & \mc{1}{802} & 73 & 195 & 554 \\ \hline
\Bowling & \mc{1}{\textbf{69}} & 10 & \mc{1}{50} & 60 & 26 & 25 \\ \hline
\Boxing & \mc{1}{\textbf{100}} & 15 &\mc{1}{ \textbf{100}} & 22 & \textbf{100} & \textbf{100} \\ \hline
\Breakout &\mc{1}{ 384} & 4 &\mc{1}{ \textbf{772}} & 39 & 1 & 364 \\ \hline
\Carnival & \mc{1}{\textbf{6372}} & 16 &\mc{1}{ 5516} & 53 & 950 & 5132 \\ \hline
\Centipede & 99207 & 39 & 94236 & 67 & \textbf{125123} & 110422 \\ \hline
\ChopperCommand & 10980 & 76 & 27220 & 73 & 1827 & \textbf{34019} \\ \hline
\CrazyClimber & 36160 & 4 & 36940 & 58 & 37110 & \textbf{98172} \\ \hline
\DemonAttack & 20116 & 33 & 16025 & 41 & 443 & \textbf{28159} \\ \hline
\DoubleDunk & -14 & 41 & 21 & 41 & -19 & \textbf{24} \\ \hline
\ElevatorAction & 13480 & 26 & 10820 & 27 & 730 & \textbf{18100} \\ \hline
\Enduro &\mc{1}{ \textbf{500}} & 66 & \mc{1}{359} & 38 & 1 & 286 \\ \hline
\FishingDerby & 30 & 39 & 6 & 62 & -92 & \textbf{38} \\ \hline
\Freeway &  \mc{1}{\textbf{31}} & 32 & \mc{1}{23} & 61 & 0 & 0 \\ \hline
\Frostbite & \mc{1}{902} & 12 &  \mc{1}{\textbf{2672}} & 38 & 137 & 271 \\ \hline
\Gopher & 18256 & 19 & 15808 & 53 & 1019 & \textbf{20560} \\ \hline
\Gravitar &\mc{1}{ 3920} & 62 &  \mc{1}{\textbf{5980}} & 62 & 395 & 2850 \\ \hline
\Hero &  \mc{1}{\textbf{12985}} & 37 & 11524 & 69 & 1324 & 12860 \\ \hline
\IceHockey &  \mc{1}{\textbf{55}} & 89 &\mc{1}{ 49} & 89 & -9 & 39 \\ \hline
\JamesBond &  \mc{1}{\textbf{23070}} & 0 & \mc{1}{10080} & 30 & 25 & 330 \\ \hline
\JourneyEscape & \mc{1}{40080} & 38 &  \mc{1}{\textbf{40600}} & 67 & 1327 & 7683 \\ \hline
\Kangaroo &  \mc{1}{\textbf{8760}} & 8 &\mc{1}{ 5320} & 31 & 90 & 1990 \\ \hline
\Krull &  \mc{1}{\textbf{6030}} & 28 & 4884 & 42 & 3089 & 5037 \\ \hline
\KungFuMaster &  \mc{1}{\textbf{63780}} & 21 & 42180 & 43 & 12127 & 48855 \\ \hline
\Montezuma & 0 & 14 &  \mc{1}{\textbf{540}} & 39 & 0 & 0 \\ \hline
\MsPacman & 21695 & 21 & 18927 & 23 & 1709 & \textbf{22336} \\ \hline
\NameThisGame & 9354 & 14 & 8304 & 25 & 5699 & \textbf{15410} \\ \hline
\Pong & \mc{1}{\textbf{21}} & 17 &\mc{1}{ \textbf{21}} & 35 & -21 & \textbf{21} \\ \hline
\Pooyan & 11225 & 8 & 10760 & 16 & 910 & \textbf{17763} \\ \hline
\PrivateEye & -99 & 18 & \mc{1}{\textbf{2544}} & 44 & 58 & 100 \\ \hline
\Qbert & 3705 & 11 & 11680 & 35 & 133 & \textbf{17343} \\ \hline
\RiverRaid & \mc{1}{\textbf{5694}} & 18 &\mc{1}{ 5062} & 37 & 2179 & 4449 \\ \hline
\RoadRunner & \mc{1}{\textbf{94940}} & 25 &  \mc{1}{68500} &  41  & 245 & 38725 \\ \hline
\RobotTank &\mc{1}{ \textbf{68}} & 34 &\mc{1}{ 52} & 34 & 2 & 50 \\ \hline
\Seaquest &\mc{1}{ \textbf{14272}} & 25 &\mc{1}{ 6138} & 33 & 288 & 5132 \\ \hline
\SpaceInvaders &\mc{1}{ 2877} & 21 & \mc{1}{\textbf{3974}} & 34 & 112 & 2718 \\ \hline
\StarGunner &\mc{1}{ 1540} & 19 & \mc{1}{\textbf{4660}} & 18 & 1345 & 1207 \\ \hline
\Tennis &\mc{1}{ \textbf{24}} & 21 & \mc{1}{\textbf{24}} & 36 & -24 & 3 \\ \hline
\TimePilot & 35000 & 9 & 36180 & 29 & 4064 & \textbf{63855} \\ \hline
\Tutankham & 172 & 15 & 204 & 34 & 64 & \textbf{226} \\ \hline
\UpNDown &\mc{1}{ \textbf{110036}} & 12 & 54820 & 14 & 746 & 74474 \\ \hline
\Venture & \mc{1}{\textbf{1200}} & 22 & \mc{1}{980} & 35 & 0 & 0 \\ \hline
\VideoPinball & \mc{1}{\textbf{388712}} & 43 & 62075 & 43 & 55567 & 254748 \\ \hline
\WizardOfWor & \mc{1}{\textbf{121060}} & 25 & 81500 & 27 & 3309 & 105500 \\ \hline
\Zaxxon & \mc{1}{\textbf{29240}} & 34 & 15680 & 31 & 0 & 22610 \\ \hline

\multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\ \hline
\# Times Best (\numgames\ games)& \multicolumn{2}{c|}{\textbf{26}} & \multicolumn{2}{c|}{13} & \multicolumn{1}{c|}{1} & \multicolumn{1}{c|}{19} \\ \hline
\# Times Better than \IW\ (\numgames\ games) & \multicolumn{2}{c|}{--} & \multicolumn{2}{c|}{16} & \multicolumn{1}{c|}{1}  & \multicolumn{1}{c|}{\textbf{19}} \\ \hline
\# Times Better than \BFS\ (\numgames\ games) & \multicolumn{2}{c|}{\textbf{34}} & \multicolumn{2}{c|}{--} & \multicolumn{1}{c|}{1}  & \multicolumn{1}{c|}{25} \\ \hline
\# Times Better than \UCT\ (\numgames\ games) & \multicolumn{2}{c|}{\textbf{31}} & \multicolumn{2}{c|}{26} & \multicolumn{1}{c|}{1}  & \multicolumn{1}{c|}{--} \\ \hline

\end{tabular}
}

 \caption{\small Performance that results form various  lookahead algorithms in \numgames\  Atari games. The algorithms, \BRFS, $\IW(1)$, \BFS, and \UCT, are evaluated
 over $10$ runs (episodes) for each game. 
%%% Averages scores  over 5 runs shown with ``random'' seeds $0$ to $4$,  
%%% this does not apply to UCT and BRFS if taken from JAIR ..
%% 
The maximum episode duration is  $18,000$  frames and every algorithm is limited to a lookahead budget of 150,000 simulated frames.
% , except for UCT limited to 500 rollouts, 300 frames deep instead. 
Figures for \BRFS\ and \UCT\ taken from \cite{bellemare:jair2013}.
Average  CPU times per action in seconds, rounded to nearest integer, shown for  $\IW(1)$ and \BFS. 
Numbers  in bold show best performer in terms of average score, while numbers shaded in light grey show scores  that are better than \UCT's. 
Bottom part of the table shows pairwise comparisons among the algorithms. 
% Entries shown as ``--'' are still running at the time of submission. Each run can take up to 24 hours of CPU time. Eventually
% all 55 games from  \cite{bellemare:jair2013}  to be included.
}
\label{performance}

\end{table*}

We tested \IW(1) and \BFS\ over \numgames\  of the $55$ games 
considered in \cite{bellemare:jair2013}, from now on  abbreviated as BNVB.\footnote{
We left out  \textsc{Skiing} as the reported figures apparently use a  different reward structure.}
%% check
% \footnote{
% Entries marked as  ``--'' in the table are   still running at the time of submission. 
% Each run can take up to $24$ hours of CPU time and there are $10$ runs per game and algorithm. 
% Eventually all of  the  $55$ games will  be included in the evaluation.}
Table~\ref{performance} shows the performance of $\IW(1)$ and \BFS\
in comparison  with breadth-first search (\BRFS) and \UCT.
Videos of selected games played by $\IW(1)$, \BFS, and \UCT\ can be 
seen in Youtube.\footnote{\url{http://www.youtube.com/playlist?list=PLXpQcXUQ_CwenUazUivhXyYvjuS6KQOI0}.}
The discount factor used by  all the algorithms is $\gamma=0.995$.
The scores reported for \BRFS\ and \UCT\ are taken from BNVB.  
Our experimental setup  follows theirs except that a maximum budget of  $150,000$ simulated frames 
is applied to $\IW(1)$, \BFS, and \UCT. \UCT\ uses this budget by running
 $500$   rollouts of depth $300$. % \footnote{In \UCT,  $500$ base policy rollouts
% of depth $300$ is equivalent to a budget of $150,000$ simulated frames.}.
The bound on the number of simulated frames is like a bound on
lookahead time, as most of the time in the lookahead  is spent
in calls to the emulator for computing the next RAM state. 
This is why the average time per action is similar to all the algorithms
except $\IW(1)$, that due to its aggressive pruning does not always use
the full budget and takes less time per action on average.

@@

\noindent \textbf{Acknowledgments.} We thank the people involved in
the creation and maintenance of the Arcade Learning Environment that
have made this work possible. This research was co-funded by the
Australian Research Council linkage grant LP11010015.
